|

### 2021-05-06 appendix/appendix.Rmd
edit text appendix

[^freire]: Independent Researcher, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

@@

In their model, every local public goods project of size $x$ generates a concave benefit $b(x)$, and there are convex costs associated with the project.\footnote{Concave means that $b'>0$ and $b''<0$. Convex means that $c'>0$ and $c''>0$. We assume that the derivatives are well defined throughout the analysis. Moreover, we drop the constituency index to make the model more intuitive, so the reader should assume that we always employ a symmetric Nash equilibrium.} The first type of cost, $c_1(x)$, comprises the expenses within the constituency (e. g., hire a local company to build the project). The second cost, $c_2(x)$, captures the expenses outside the constituency (e. g., hire a company from another state). The third cost $c_3(x)$, captures the externalities generated by the project (e. g., how much prices shift because local economic factors are being used to provide the project). The total cost is equal to $c(x) = c_1(x) + c_2(x) + c_3(x)$. The tax burden generated by the project is equal to $T(x) = c_1(x) + c_2(x)$.

Following their model, projects are economically efficient when the marginal costs are equal the marginal benefits of increase the project size. This leads to the economic optimal project size, $x^E$. The optimal project size is $b'(x^E) - c_1'(x^E) - c_2'(x^E) - c_3'(x^E) = 0$. However, the projects that are actually implemented have a different structure. First, assume that the constituency in question has a tax burden $t = 1/n$, where $n$ represents the number of constituencies. Also, suppose that benefits are distorted by the fact that costs within the constituency ($c_1(x)$) become investments in local firms. Therefore, the costs and benefits of implementing a project with size $x$ have the following structure:

@@

And rearranging the terms, yields to the following expression for $\dfrac{dx}{dn}$:

@@

Note that the numerator is always positive, as the marginal investment costs inside and outside the district increases with project size. The _law of 1/n_ holds when the denominator is negative. $b''(x) < 0$ by assumption, and $-n^{-1}[c_1''(x) + c_2''(x)] - c_3''(x) < 0$ as $c_1'', c_2'', c_3'' > 0$. Then, the _law of 1/n_ is true when $c_1''(x)$ is smaller than $n^{-1}[c_1''(x) + c_2''(x)] + c_3''(x) - b''(x)$.

As the above condition suggests, the _law of 1/n_ holds only in specific situations. @primo2008distributive advances the theory by considering other situations where the law may not hold, and also argues that there are cases in which a reverse _law of 1/n_ holds, i. e., larger legislatures leading to lower expenditures.

Following @primo2008distributive, let $n$ be the number of districts, $m$ the number of citizens in each district, and $nm$ is the total population in the country. Consider a local public good that generates a per capita benefit according to size $x$ of $b(x, m) = x^\alpha m^{\beta-1}$, where $\beta$ is the degree of congestion of the public good, i. e. how much the addition of individuals reduce the benefits for other individuals (note that the lower the $\beta$, the higher the congestion). In terms of costs, consider a linear cost function $C(x) = x$, and in terms of taxation, assume that the people in the district share taxes between their payments and the central government's. The degree that taxes are shared is denoted by $s$. Moreover, there is a deadweight loss of the taxes $\theta \geq 1$. The tax then becomes: $t = \left(\dfrac{(n – ns + s)x + (ns – s)X}{nm}\right)^\theta$. Collecting terms, the citizens receive the following net benefit of a project with size $x$:

@@

Therefore, we see that both the _law of 1/n_ and its reverse formulation are equally plausible. Here we show how the results change just by changing the levels of deadweight losses, but authors have suggested other reasons why the _law of 1/n_ may not apply, such as bicameralism [@chen_malhotra_2007], popular initiatives [@matsusaka2005endogeneity], type of government [@coate2011government], supermajorities [@lee2015supermajority], political fragmentation [@lledo2003electoral], and ideology [@bjedov2014impact].

@@

The first step in our systematic review consisted in gathering a study sample. We started our data collection with a manual search based on a set of keywords we scouted from the distributive politics literature. This search produced a database with many entries that were unrelated to our subject of investigation. To reduce the number of false positives in our sample, we restricted our search to studies that cited Weingast, Shepsle and Johnsen's 1981 paper "_The Political Economy of Benefits and Costs: A Neoclassical Approach to Distributive Politics_", which is a seminal contribution to the field. Although \href{https://scholar.google.com/}{Google Scholar} reports the article has received \href{https://scholar.google.com/scholar?um=1&ie=UTF-8&lr&cites=13117579863846712459}{2,180} citations, our search resulted in 2,664 records on the 21^st^ of November 2019.

We webscraped three large academic databases: \href{https://scholar.google.com/}{Google Scholar} (n = 1001); \href{https://academic.microsoft.com/home}{Microsoft Academic} (n = 927); and \href{https://www.scopus.com/}{Scopus} (n = 736). The `R` script we wrote extracted the article title, abstract, authors, year, journal of publication, and database from which the record originated. Our code is available in section \ref{sub:scrap} below. We screened these results with an English language and article restriction, that is, we excluded all records written in other languages and all that were not academic papers, such as book chapters or doctoral theses. We set no restriction to unpublished articles.

We complemented this by doing a term search on Google Scholar after having finished the entire eligibility procedure on the first sample. We formulated the search string based on the terms and expressions that appeared most commonly in the articles included in our meta-analysis. The choice of string translates the central point of our inquiry: the relationship between legislature size and public spending. The search string was as follows: `("upper chamber size" OR "lower chamber size" OR "council size" OR "parliament size" OR "legislature size" OR "number of legislators" OR "legislative size") AND ("spending" OR "expenditure" OR "government size")`. We scraped Google Scholar on March 5th, 2021. Using the titles of this new database's records, we did a fuzzy matching with the database of @weingast1981political citations mentioned previously. We then performed the exact same eligibility procedures, further explained below. This resulted in 3,041 additional records.

Combining the two search strategies, we assessed a total of 5,705 records.

@@

The third phase consisted of filling out tables for each of the remaining 47 articles to systematically evaluate their eligibility. Since authors use different measures for government spending and the number of lower/upper house members, we extracted all coefficients that provided this information. In this phase, we also collected information on whether the paper had been published, and if it explicitly discussed the \textit{law of 1/n}. Upon choosing the variables, we excluded the non-conforming studies, arriving at our final sample of `r length(unique(dat$authoryear))` articles.

@@

Our meta-analytic data are comprised of two datasets. The first dataset has the main coefficients reported in the studies. These data include only the most rigorous model from each paper, that is, those estimated with the largest sample size, most control variables, and fixed effects if the authors added them. If the article employed a regression discontinuity design, we chose the coefficient from the optimal bandwidth or from the intermediate one. This sample encompasses `r dim(dat)[1]` estimates, as `r as.numeric(table(table(dat$id))[2])` articles analysed two dependent or independent variables of interest. Our second sample, in contrast, contains all the `r dim(fulldat)[1]` effect sizes reported in the `r length(unique(dat$id))` papers.

@@

Studies were included in our sample regardless of their publication status. From the `r length(unique(dat$id))` papers in the sample, `r as.numeric(table(unique(select(dat, id, published))$published)[1])` were published while `r as.numeric(table(unique(select(dat, id, published))$published)[2])` were not published.

@@

Our sample differs considerably in regards to research design. One remarkable difference is that several authors apply the logics of the *law of 1/n*, which was built with majoritarian systems in mind, to non-majoritarian democracies. In the sample, `r as.numeric(table(unique(select(dat, id, elecsys2))$elecsys2)[1])` of the papers study *Majoritarian* systems while `r as.numeric(table(unique(select(dat, id, elecsys2))$elecsys2)[2])` study *Non-Majoritarian* electoral systems.\footnote{Note that for the \textit{law of 1/n} to be valid in a non-majoritarian system, we need to assume that despite the fact that politicians are able to campaign in every place in the district, the votes are geographically concentrated. The concentration facilitates politicians to use pork-barrel projects to captivate their electoral supporters.}

@@

-   `r as.numeric(table(unique(select(dat, id, depvar2))$depvar2)[1])` Per Capita Expenditure papers
-   `r as.numeric(table(unique(select(dat, id, depvar2))$depvar2)[3])` Natural Log of Per Capita Expenditure papers
-   `r as.numeric(table(unique(select(dat, id, depvar2))$depvar2)[2])` Expenditure as a Percentage of the GDP papers

@@

Most papers in our sample analyse the number of legislators in the lower house (`r as.numeric(table(dat$indepvar2)[2])`). The second most frequent independent variable is the number of legislators in the upper house (`r as.numeric(table(dat$indepvar2)[1])`). Finally, the minority of papers use the natural log of the number of legislators in the lower house as an independent variable (`r as.numeric(table(dat$indepvar2)[3])`). As we noted above, some papers had multiple coefficients, and thus the total number of coefficients is `r dim(dat)[1]`, while the number of papers is only `r length(unique(dat$id))`.

```{r, fig.width=6, fig.height=3, fig.cap="Independent variables across the law of 1/n studies", size="footnotesize", echo = FALSE, cache=TRUE}
dat %>%
  select(id, indepvar2) %>%
  unique() %>%
  mutate(indepvar2 = factor(indepvar2, labels = c("K (upper house legislators)",
                                                  "N (lower house legislators)",
                                                  "Log N"))) %>%
  ggplot(aes(x = indepvar2)) +
    geom_bar(color = "black") +
  labs(x = "",
       y = "Independent Variables") +
  coord_flip() +
  theme_bw()
```

@@

The coefficients in the papers vary considerably. In this section, we plot a histogram of the coefficients for all measurements included in the meta-analytic dataset.

@@

In this paper, we employ a random-effects model. The empirical papers testing the *law of 1/n* are very diverse. We tried to capture some of this diversity by considering the main dependent and independent variables separately, but they have at least three other important sources of dispersion:

1.  **Subjects**: Counties, Municipalities, States, Provinces, Countries.
2.  **Electoral systems**: Majoritarian, PR, Mixed.
3.  **Modelling strategies**: Panel data, Standard OLS, IV, RDD.

@@

Another crucial assumption in meta-analysis is that the coefficients should be independent [@harrer2019doing; @cheung2019guide; @veroniki2016methods; @borenstein2011introduction]. This assumption states that for our findings to be consistent, the coefficients must come from different sources of variation. However, in the political economy literature, authors frequently use similar datasets, and almost all papers fit more than one model for the same variable. While our restricted dataset contains `r nrow(dat)` estimates, our full dataset has `r nrow(fulldat)` coefficients. This is because the papers report an average of `r round(mean(as.numeric(table(fulldat$id))), 2)` coefficients. To correct for the violation of the independent assumption, we use a multi-level random effects model [@cheung2014modeling]. We add two extra levels to the regular random effects model, one that indicates the publication ID and another that indicates the data used in the original article. These levels are assumed to remove dependence structures in the data, therefore improving the estimates of our coefficient of interest, which is the effect of legislature size on public spending.

@@

| Publication IDs                             | Source of Dependence                                                     |
|---------------------------------------------|--------------------------------------------------------------------------|
| 3, 42, 132, 165, 439, 441, 467, 505         | US States Data                                                           |
| 408, 208, A258                              | US Municipalities Data                                                   |
| 849, 578                                    | US Municipalities Data and Same Author in Two Different Studies          |

All the remaining papers receive a unique index. As the number of the papers with these dependencies are not very high, and many characteristics within papers vary (e. g., state-wise ad-hoc exclusions, start and end points variations, region selection, etc), we show that it makes little difference to change from this index to the paper IDs. In regards to the full dataset, the results change considerably, because authors fit several models within the same paper. In any case, we use multilevel random effects for all the estimated models in this paper.

@@

```{r, fig.width=8, fig.height=10, fig.cap="Effect of log lower houses size (logN) on the log of per capita government expenditure (logExpPC)", echo = FALSE, warning=FALSE, cache=TRUE}
build_forest(mod, NULL)
```

@@

```{r, fig.width=8, fig.height=5, fig.cap="Effect of upper house size (K) on the public expenditure share of the GDP (PCTGDP)",warning=FALSE, cache=TRUE, echo = FALSE}
build_forest(mod, NULL)
```

@@

1. In terms of the modeling, passing from `OLS` to `PANEL` or `RDD` decreases the detected effects.

@@

4. Compared with OLS, Panel and RDD report significantly smaller coefficients.

>>>>

[^freire]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

@@

In their model, every local public goods project of size $x$ generates a concave benefit $b(x)$, and there are convex costs associated with the project.\footnote{Concave means that $b'>0$ and $b''<0$. Convex means that $c'>0$ and $c''>0$. We assume that the derivatives are well defined throughout the analysis. Moreover, we drop the constituency index to make the model more intuitive, so the reader should assume that we always employ a symmetric Nash equilibrium.} The first type of cost, $c_1(x)$, comprises the expenses within the constituency (e. g., hiring a local company for the project). The second type of cost, $c_2(x)$, captures the expenses outside the constituency (e. g., hiring a company from another state). Finally, the third cost, $c_3(x)$, captures the externalities generated by the project (e. g., how much prices shift because local economic factors are being used to provide the project). The total cost is equal to $c(x) = c_1(x) + c_2(x) + c_3(x)$. The tax burden generated by the project is equal to $T(x) = c_1(x) + c_2(x)$.

Projects are economically efficient when the marginal costs are equal to the marginal benefits of the project size. This leads to the economic optimal project size, $x^E$. The optimal project size is $b'(x^E) - c_1'(x^E) - c_2'(x^E) - c_3'(x^E) = 0$. However, the projects that are actually implemented have a different structure. First, assume that the constituency in question has a tax burden $t = 1/n$, where $n$ represents the number of constituencies. Also, suppose that benefits are distorted by the fact that costs within the constituency ($c_1(x)$) become investments in local firms. Therefore, the costs and benefits of implementing a project with size $x$ have the following structure:

@@

And rearranging the terms, we find the following expression for $\dfrac{dx}{dn}$:

@@

Note that the numerator is always positive, as the marginal investment costs inside and outside the district increase according to the project size. The _law of 1/n_ holds when the denominator is negative. $b''(x) < 0$ by assumption, and $-n^{-1}[c_1''(x) + c_2''(x)] - c_3''(x) < 0$ as $c_1'', c_2'', c_3'' > 0$. Then, the _law of 1/n_ is true when $c_1''(x)$ is smaller than $n^{-1}[c_1''(x) + c_2''(x)] + c_3''(x) - b''(x)$.

As the above condition suggests, the _law of 1/n_ holds only in specific situations. @primo2008distributive advance that theory by considering other situations where the _law_ may not hold, and they also argue that there are cases in which a _reverse law of 1/n_ holds, that is, larger legislatures may lead to lower public expenditures.

Following @primo2008distributive, let $n$ be the number of districts, $m$ the number of citizens in each district, and $nm$ is the total population in the country. Consider a local public good that generates a per capita benefit of $b(x, m) = x^\alpha m^{\beta-1}$ according to size $x$, where $\beta$ is the degree of congestion of the public good, that is, how much the addition of individuals reduce the benefits for other individuals (note that the lower the $\beta$, the higher the congestion). In terms of costs, consider a linear cost function $C(x) = x$, and in terms of taxation, assume that the people in the district pay both local and federal government taxes. The degree that taxes are shared is denoted by $s$. Moreover, there is a deadweight loss of the taxes $\theta \geq 1$. The tax then becomes: $t = \left(\dfrac{(n – ns + s)x + (ns – s)X}{nm}\right)^\theta$. Therefore, the citizens receive the following net benefit of a project with size $x$:

@@

Thus, we see that both the _law of 1/n_ and its reverse formulation are equally plausible. Here we show how the results change just by changing the levels of deadweight losses, but authors have suggested other reasons why the _law of 1/n_ may not apply, such as bicameralism [@chen_malhotra_2007], popular initiatives [@matsusaka2005endogeneity], type of government [@coate2011government], supermajorities [@lee2015supermajority], political fragmentation [@lledo2003electoral], and ideology [@bjedov2014impact].

@@

The first step in our systematic review consisted in gathering a study sample. We started our data collection with a manual search based on a set of keywords we scouted from the distributive politics literature. This search produced a database with many entries that were unrelated to our subject of investigation. To reduce the number of false positives in our sample, we restricted our search to studies that cited Weingast, Shepsle and Johnsen's 1981 paper "_The Political Economy of Benefits and Costs: A Neoclassical Approach to Distributive Politics_", which is the fundamental contribution to the field. Although \href{https://scholar.google.com/}{Google Scholar} reports the article has received \href{https://scholar.google.com/scholar?um=1&ie=UTF-8&lr&cites=13117579863846712459}{2,180} citations, our search resulted in 2,664 records on the 21^st^ of November 2019.

We webscraped three large academic databases: \href{https://scholar.google.com/}{Google Scholar} (n = 1001); \href{https://academic.microsoft.com/home}{Microsoft Academic} (n = 927); and \href{https://www.scopus.com/}{Scopus} (n = 736). The `R` script we wrote extracted the article title, abstract, authors, year, journal of publication, and database from which the record originated. Our code is available in the `R` script available in this repository. We screened these results with an English language and article restriction, that is, we excluded all records written in other languages and all that were not academic papers, such as book chapters or doctoral theses. We set no restriction to unpublished articles.

We complemented this process by doing a term search on Google Scholar after having finished the entire eligibility procedure in the first sample. We formulated the search string based on the terms and expressions that appeared most frequently in the articles included in our meta-analysis. The choice of string translates the central point of our inquiry: the relationship between legislature size and public spending. The search string was as follows: `("upper chamber size" OR "lower chamber size" OR "council size" OR "parliament size" OR "legislature size" OR "number of legislators" OR "legislative size") AND ("spending" OR "expenditure" OR "government size")`. We scraped Google Scholar on March 5^th^, 2021. Using the titles of this new database's records, we did a fuzzy matching with the database of @weingast1981political citations mentioned previously. We then performed the exact same eligibility procedures, further explained below. This resulted in 3,041 additional records. Combining the two search strategies, we assessed a total of 5,705 records.

@@

The third phase consisted of filling out tables for each of the remaining 47
articles to systematically evaluate their eligibility. Since authors use
different measures for government spending and the number of lower/upper house
members, we extracted all coefficients that provided this information. In this
phase, we also collected information on whether the paper had been published,
and if it explicitly discussed the _law of 1/n_. Upon choosing the variables,
we excluded two studies from the first sample, as they did not have the
dependent/independent variables we collected for our meta-analysis. We then
included the 3 papers we found in the second search, as both conformed to our
criteria, and compiled our final sample of `r length(unique(dat$authoryear))`
articles.

@@

Our meta-analytic data are comprised of two datasets. The first dataset has the main coefficients reported in the selected studies. These data include only the most rigorous model from each paper, that is, those estimated with the largest sample size, most control variables, and fixed effects if the authors added them. If the article employed a regression discontinuity design, we chose the coefficient from the optimal bandwidth or from the intermediate one. This sample encompasses `r dim(dat)[1]` estimates, as `r as.numeric(table(table(dat$id))[2])` articles analysed two dependent or independent variables of interest. Our second sample, in contrast, contains all the `r dim(fulldat)[1]` effect sizes reported in the `r length(unique(dat$id))` papers.

@@

Studies were included in our sample regardless of their publication status. From the `r length(unique(dat$id))` papers in the sample, `r as.numeric(table(unique(select(dat, id, published))$published)[2])` were published while `r as.numeric(table(unique(select(dat, id, published))$published)[1])` were not published.

@@

Our sample differs considerably in regards to research design. One remarkable difference is that several authors apply the logics of the *law of 1/n*, which was built with majoritarian systems in mind, to non-majoritarian democracies. In the sample, `r as.numeric(table(unique(select(dat, id, elecsys2))$elecsys2)[1])` of the papers study *Majoritarian* systems while `r as.numeric(table(unique(select(dat, id, elecsys2))$elecsys2)[2])` study *Non-Majoritarian* electoral systems.\footnote{Note that for the \textit{law of 1/n} to be valid in a non-majoritarian system, we need to assume that despite the fact that politicians are able to campaign in every place in the district, votes are geographically concentrated. The concentration facilitates politicians to use pork-barrel projects to captivate their electoral supporters.}

@@

-   `r as.numeric(table(unique(select(dat, id, depvar2))$depvar2)[1])` Per Capita Expenditure papers
-   `r as.numeric(table(unique(select(dat, id, depvar2))$depvar2)[2])` Expenditure as a Percentage of the GDP papers
-   `r as.numeric(table(unique(select(dat, id, depvar2))$depvar2)[3])` Natural Log of Per Capita Expenditure papers

@@

Most papers in our sample analyse the number of legislators in the lower house (`r as.numeric(table(dat$indepvar2)[3])`). The second most frequent independent variable is the number of legislators in the upper house (`r as.numeric(table(dat$indepvar2)[1])`). Finally, the minority of papers use the natural log of the number of legislators in the lower house as an independent variable (`r as.numeric(table(dat$indepvar2)[1])`). As we noted above, some papers had multiple coefficients, and thus the total number of coefficients is `r dim(dat)[1]`, while the number of papers is only `r length(unique(dat$id))`.

```{r, fig.width=6, fig.height=3, fig.cap="Independent variables across the law of 1/n studies", size="footnotesize", echo = FALSE, cache=TRUE}
dat %>%
  select(id, indepvar2) %>%
  unique() %>%
  mutate(indepvar2 = factor(indepvar2, labels = c("K (upper house legislators)",
                                                  "Log N",
                                                  "N (lower house legislators)"
                                                  ))) %>%
  ggplot(aes(x = indepvar2)) +
    geom_bar(color = "black") +
  labs(x = "",
       y = "Independent Variables") +
  coord_flip() +
  theme_bw()
```

@@

The coefficients in the papers vary considerably. We plot a histogram of the coefficients for all measurements included in the meta-analytic dataset. Most coefficients and standard deviations are close to zero.

@@

In this paper, we employ a random effects model. The empirical papers testing the *law of 1/n* are very diverse. We tried to capture some of this diversity by considering the main dependent and independent variables separately, but they have at least three other important sources of dispersion:

1.  **Study sample**: Counties, Municipalities, States, Provinces, Countries.
2.  **Electoral systems**: Majoritarian, PR, Mixed.
3.  **Modelling strategies**: Panel data, Standard OLS, IV, RDD.

@@

Another crucial assumption in meta-analysis is that the coefficients should be independent [@harrer2019doing; @cheung2019guide; @veroniki2016methods; @borenstein2011introduction]. This assumption states that for our findings to be consistent, the coefficients must come from different sources of variation. However, in the political economy literature, authors frequently use similar datasets, and almost all papers fit more than one model with similar variables. While our restricted dataset contains `r nrow(dat)` estimates, our full dataset has `r nrow(fulldat)` coefficients. This is because the papers report an average of `r round(mean(as.numeric(table(fulldat$id))), 2)` coefficients. To correct for the violation of the independent assumption, we use a multilevel random effects model [@cheung2014modeling]. We add two extra levels to the regular random effects model, one that indicates the publication ID and another that indicates the data used in the original article. These levels are assumed to remove dependence structures in the data, therefore improving the estimates of our coefficient of interest, which is the effect of legislature size on public spending.

@@

| Publication ID                              | Source of Dependence                                                     |
|---------------------------------------------|--------------------------------------------------------------------------|
| 3, 42, 132, 165, 439, 441, 467, 505         | US States Data                                                           |
| 408, 208, A258                              | US Municipalities Data                                                   |
| 849, 578                                    | US Municipalities Data and Same Author in Two Different Studies          |

All the remaining papers received a unique index. As the number of the papers with these dependencies is not very high, and many characteristics within papers vary (e. g., state-wise ad-hoc exclusions, start and end points variations, region selection, etc), we show that it makes little difference to change from this index to the paper IDs. In regard to the full dataset, the results change considerably, because authors fit several models within the same paper. In any case, we use multilevel random effects for all the estimated models in this paper.

@@

```{r, fig.width=6, fig.height=4, fig.cap="Effect of log lower houses size (logN) on the log of per capita government expenditure (logExpPC)", echo = FALSE, warning=FALSE, cache=TRUE}
build_forest(mod, NULL)
```

@@

```{r, fig.width=8, fig.height=4, fig.cap="Effect of upper house size (K) on the public expenditure share of the GDP (PCTGDP)",warning=FALSE, cache=TRUE, echo = FALSE}
build_forest(mod, NULL)
```

@@

1. In regard to the estimation method, papers that use `PANEL` data or `RDD`
   designs have smaller effects than articles that employ `OLS` estimations.

@@

4. Compared with `OLS`, `PANEL` and `RDD` report significantly smaller coefficients.
